---
title: "Budget Report"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    self_contained: FALSE
    lib_dir: libs  
    mathjax: "default"
    toc: NO
    includes:
      after_body: footer.html
---

<style> 
  #header { 
    background: #2274cc; /* Old browsers */
    background: -moz-linear-gradient(left, #2274cc 0%, #26c5d3 36%, #61bf61 100%); /* FF3.6-15 */
    background: -webkit-linear-gradient(left, #2274cc 0%,#26c5d3 36%,#61bf61 100%); /* Chrome10-25,Safari5.1-6 */
    background: linear-gradient(to right, #2274cc 0%,#26c5d3 36%,#61bf61 100%); /* W3C, IE10+, FF16+, Chrome26+,  Opera12+, Safari7+ */
    filter: progid:DXImageTransform.Microsoft.gradient( startColorstr='#2274cc', endColorstr='#61bf61',GradientType=1 ); /* IE6-9 */
    color: white;
    height: 100px;
    display:flex;
    align-items: center;
    justify-content: center;
  }

  h1.title {
    margin: auto;
  }
</style>

<br>  

#### Expense report for home budgeting.


```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = (normalizePath(getwd())) )

knitr::opts_chunk$set(fig.width = 8, fig.asp = 1/2)
require('magrittr') # load for "%>%" used to format DT tables

```

```{r setData_Func, include=TRUE, echo=FALSE}
  # load data
   creditCardHistory <- data.table::copy(homeBudget::creditCardExample)
```

```{r computeSummaries, include=FALSE}
 
  # bucket description of credit card transactions into budget categories
  creditCardHistory[ , ':='(category = homeBudget::addCategories(Description = Description,
                                    myCategories = homeBudget::categoryPatternsExample))]
  
  # summary of the continuous values for use in data.table.
  summaryDT <- function(x, digits = NULL, suffix = NULL){
    N <- function(x) length(x)
    q25 <- function(x) quantile(x, 0.25,names = FALSE, type = 2)
    q50 <- function(x) quantile(x, 0.5,names = FALSE, type = 2)
    q75 <- function(x) quantile(x, 0.75,names = FALSE, type = 2)
    stats <- c(N, min, q25, q50, mean, sd, q75, max)
    results <- lapply(stats, function(f,x){f(x)},x)
    if(!is.null(digits)){
      results <- lapply(results, function(x){round(x, digits)})
    }
    names(results) = paste0(c("N", "min","q25","q50","mean","sd","q75","max"), 
                            seperator = suffix)
    results
  }
  
  # computing rolling sums of ordered observations.
  # This functions computes the rolling sums for the previous and including 'k' days.
  rollSumTime <- function(x, time, k){
    if(anyDuplicated(time) > 0){
      stop("Duplicates in `time` not allowed")
    }
    unlist(Map(function(x, time, k, timeList){
      d <- as.numeric(timeList) - as.numeric(time)
      sum(x[d <= 0 & d > -k])
    }, list(x), time, list(k), list(timeList = time)))
  }
  
  # For days with no spending append a record to the data-set for an amount spent of $0.
  appendDaysAbsent <- function(creditCardHistory,
                             minDate = NULL,
                             maxDate = NULL,
                             category = NULL){
    if(is.null(minDate)) minDate <- min(creditCardHistory$PostingDate)
    if(is.null(maxDate)) maxDate <- max(creditCardHistory$PostingDate)
    padDates <- setdiff(seq.Date(minDate, maxDate, by = 1),
                        creditCardHistory$PostingDate)
    if(length(padDates) > 0){
      allDatesDT <- data.table::data.table("PostingDate" = padDates,
                                           "Amount" = 0)
      if(!is.null(category)) allDatesDT[ , ':='(category = category)]
    } else return(creditCardHistory)
    data.table::rbindlist(l = list(creditCardHistory, allDatesDT), 
                          use.names = TRUE, fill = TRUE)
  }
  
  creditCardHistoryPad <- appendDaysAbsent(creditCardHistory)
  creditCardHistoryPad[, ':='(category = "Total")]
  
  creditCardHistoryPad <-
    data.table::rbindlist(l = list(creditCardHistoryPad,
                data.table::rbindlist(l = 
                          lapply(X = unique(creditCardHistory$category), 
         FUN = function(x){appendDaysAbsent(creditCardHistory[x==category], 
                                            minDate = min(creditCardHistory$PostingDate),
                                            maxDate = max(creditCardHistory$PostingDate),
                                            category = x)
           }),
         use.names = TRUE, fill = TRUE)))
  
  # rolling sum data
  rollingSumData <- creditCardHistoryPad[,.(Amount = sum(Amount)),
                                      by = c("category", "PostingDate")
                                      ][ ,":="(rollAmount= rollSumTime(x = Amount,
                                           time = PostingDate,
                                           k = 28),
                        minDate = min(PostingDate),
                        minN = .N), by = "category"][(minDate + 28 < PostingDate |
                                                        minN ==1), ]
  

    # Summary of monthly data
    monthRollSumSummary <- rollingSumData[ , append(summaryDT(x = rollAmount,
                                                                   digits = 1), 
                                            list("N_records" = sum(Amount != 0))),
                                   by = "category"
                                   ][rev(order(q50)), ]
  
  # move the days with expenses 'N_records' for reading clarity of printed tables
  mycolnames <- names(monthRollSumSummary)
  data.table::setcolorder(x = monthRollSumSummary,
                          neworder = c(mycolnames[1:2],
                                       "N_records",
                                       mycolnames[3:(length(mycolnames)-1)]))
  mycolnames <- NULL
```


<script>
$(document).ready(function() {
  $('#myDiv').appendTo('#first-tab-sec');
});
</script>

<div id="myDiv">
<br>
<br>
<br>
```{r, include = TRUE, echo = FALSE}

 ggplot2::ggplot(rollingSumData, ggplot2::aes(x = reorder(category, rollAmount),
                                  y = rollAmount, fill = category)) + 
    ggplot2::geom_violin() +
    ggplot2::labs(x = "Date", y = "Dollars",
                title = "Density Plot of Monthly Spending") +
    ggplot2::theme(legend.position="none")
```
  
```{r roll30Summary, include=TRUE, echo=FALSE}
DT::datatable(monthRollSumSummary[ , .SD, 
                                 .SDcols = setdiff(names(monthRollSumSummary),
                                                   c("N", "category"))], 
                colnames = c('N days' = 'N_records',
                             "Std. Dev." = "sd"),
                caption = "Monthly Spending",
                rownames = as.character(monthRollSumSummary$category),
                options = list(pageLength = nrow(monthRollSumSummary)))
```
</div>
  
# Expenses {.tabset}  

## report {.tabset .tabset-fade #first-tab-sec}  
```{r expensePlotTotal, include=TRUE, echo=FALSE}
  ggplot2::ggplot(data = rollingSumData[category=="Total"],
                ggplot2::aes(x=PostingDate, y=rollAmount)) +
    ggplot2::labs(x = "Date", y = "Dollars",
                  title = "Total Spending", subtitle = "28 day rolling sum",
                  caption = "Highlighted area contains 50% of monthly spending & dashed line is the median") +
    ggplot2::geom_point()  +
    ggplot2::geom_hline(yintercept = monthRollSumSummary[category=="Total"]$q50,
                        linetype = 2, size = 1,
                        colour="red") +
      ggplot2::annotate("rect", fill = "red", alpha = 0.5,
                      xmin = as.Date(-Inf, origin = Sys.Date()),
                      xmax = as.Date(Inf, origin = Sys.Date()),
                      ymin = monthRollSumSummary[category=="Total"]$q25,
                      ymax = monthRollSumSummary[category=="Total"]$q75)

```
  
### Daily   
```{r expensePlot, include=TRUE, echo=FALSE}
ggplot2::ggplot(creditCardHistory[,.(Amount = sum(Amount)),
                                      by = c("category", "PostingDate")
                                      ],
                ggplot2::aes(x=PostingDate, y=Amount)) +
    ggplot2::labs(x = "Date", y = "Dollars",
                  title = "Expenses per day", 
                  subtitle = "Spending by Category") +
    ggplot2::geom_point()+ggplot2::facet_wrap(~category, scales = "free_y", 
                                              ncol = 3) +
    ggplot2::scale_x_date(labels = scales::date_format("%m-%Y")) +
    ggplot2::theme(axis.text.x = ggplot2::element_text(angle=30, hjust=1))

```


### Rolling 
```{r expensePlotRoll, include=TRUE, echo=FALSE}
ggplot2::ggplot(rollingSumData[!(category=="Total" | Amount==0)],
                ggplot2::aes(x = PostingDate, y = rollAmount)) +
    ggplot2::labs(x = "Date", y = "Dollars",
                  title = "28 day rolling sum", 
                  subtitle = "Spending by Category") +
    ggplot2::geom_point()+ggplot2::facet_wrap(~category, scales = "free_y", 
                                              ncol = 3) +
    ggplot2::scale_x_date(labels = scales::date_format("%m-%Y")) +
    ggplot2::theme(axis.text.x = ggplot2::element_text(angle=30, hjust=1))

```
  


## raw data
```{r expenseTable, include=TRUE, echo=FALSE}
 data.table::setorderv(creditCardHistory, "PostingDate")
 DT::datatable(creditCardHistory[ , .(PostingDate, Amount, category, Description)])

```

#

<br>
<br>


