---
title: "Income Statement"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    self_contained: FALSE
    lib_dir: libs  
    mathjax: "default"
    toc: NO
    includes:
      after_body: footer.html
---

<style> 
  #header { 
    background: #2274cc; /* Old browsers */
    background: -moz-linear-gradient(left, #2274cc 0%, #26c5d3 36%, #61bf61 100%); /* FF3.6-15 */
    background: -webkit-linear-gradient(left, #2274cc 0%,#26c5d3 36%,#61bf61 100%); /* Chrome10-25,Safari5.1-6 */
    background: linear-gradient(to right, #2274cc 0%,#26c5d3 36%,#61bf61 100%); /* W3C, IE10+, FF16+, Chrome26+,  Opera12+, Safari7+ */
    filter: progid:DXImageTransform.Microsoft.gradient( startColorstr='#2274cc', endColorstr='#61bf61',GradientType=1 ); /* IE6-9 */
    color: white;
    height: 100px;
    display:flex;
    align-items: center;
    justify-content: center;
  }

  h1.title {
    margin: auto;
  }
</style>

<br>  

#### Income Statement

The income statement (or profit & loss) measures the profitability of the home by comparing the income to the associated costs.

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = (normalizePath(getwd())) )

knitr::opts_chunk$set(fig.width = 10, fig.asp = 1/2)
require('magrittr') # load for "%>%" used to format DT tables

ggplot2::theme_set(ggthemes::theme_clean())

# time frame of report, reference point for setting the quarterly periods
  cutoff <- Sys.Date()
  
  # Print P&L table for each period
  incomeTable <- function(mytable, myperiod){
    
    mytablePeriod <- mytable[period==myperiod]
    
    if(nrow(mytablePeriod) == 0) return("No Data")
    
    tableTitle <- sprintf("P&L Quarter %i, %s to %s", 
                      unique(mytablePeriod$period), 
                      unique(mytablePeriod$origin_period),
                      unique(mytablePeriod$cutoff_period))

  incomeLength <- nrow(incomeDataSum[period==myperiod])
  costLength <- nrow(costDataSum[period==myperiod])
  
  knitr::kable(mytablePeriod[ , .(category, Amount)], 
               align = c('l', 'r'),
               col.names = NULL, 
               format = "html", 
               caption = tableTitle) %>%
    kableExtra::pack_rows("Revenue", start_row = 1, 
                          end_row = incomeLength )  %>%
    kableExtra::row_spec(row = incomeLength,
                         bold = FALSE, 
                     underline = TRUE)  %>%
    kableExtra::add_indent(positions = incomeLength)  %>%
    kableExtra::pack_rows("Expenses", 
                          start_row = incomeLength+1, 
                          end_row = (incomeLength+1)+
                            (costLength-1)) %>%
    kableExtra::row_spec(row = (incomeLength+1)+
                            (costLength-1), bold = FALSE, 
                         underline = TRUE, )  %>%
    kableExtra::add_indent(positions = (incomeLength+1)+
                            (costLength-1))  %>%
    kableExtra::pack_rows("", 
                          start_row = nrow(mytablePeriod), 
                          end_row = nrow(mytablePeriod)) %>%
    kableExtra::row_spec(row = nrow(mytablePeriod), bold = TRUE, 
                         underline = TRUE)
  }
```

```{r buildIncomeData, include=TRUE, echo=FALSE}

  # number of overlapping days between the period
  # between `x_start` to `x_End` and that from
  # `origin` to `cutoff`.
   overlap <- function(x_start, x_End, origin, cutoff){
     min_end <- pmin(cutoff, x_End)
     min_start <- pmax(origin, x_start)
     pmax(min_end - min_start + 1, 0)
   }
   
   # calculate the proportion of the income has been earned
   #  by comparing the overlap with the date range specified
   # by `orgin` and `cutoff`.
   overlapAmount <- function(dt, origin, cutoff){
     mycols <- setdiff(names(dt), "Amount")
     # drop rows that do not overlap the time window
     dt[ , c(list("Amount" = Amount * as.numeric(overlap(DateStart, DateEnd, 
                                                      origin, cutoff)) / 
                          as.numeric(DateEnd - DateStart)),
             .SD), .SDcols = mycols][Amount > 0]
   }
   
   # date range for each quarter with last quarter ending at cutoff.
   quarterly_periods <- function(cutoff){
     list(cutoff_period = cutoff - lubridate::days(c(0:3)*91),
          origin_period = cutoff - lubridate::days((c(1:4)*91)-1),
          period = 4:1)
   }
   
   # income earned per quarter
   quarterly_income <- function(dt, cutoff){
     myquarterly_periods <- quarterly_periods(cutoff)

     datalist = list()
     for(i in 1:4){
       dat <- overlapAmount(dt, myquarterly_periods$origin_period[i],
                            myquarterly_periods$cutoff_period[i])
       dat$period <- myquarterly_periods$period[i]
       dat$origin_period <- myquarterly_periods$origin_period[i]
       dat$cutoff_period <- myquarterly_periods$cutoff_period[i]
       datalist[[i]] <- dat 
     }
     data.table::rbindlist(l = datalist)
   }
    
   incomeDataPeriods <- quarterly_income(homeBudget::incomeExample, cutoff)
```


```{r buildCostData, include=TRUE, echo=FALSE}

  # load data
   costData <- data.table::copy(homeBudget::creditCardExample)

  costData[ , ':='(category = homeBudget::addCategories(Description = Description,
                              myCategories = homeBudget::categoryPatternsExample))]

  # add information about the quarter in which the cost transaction occured
  timePeriods <- data.table::as.data.table(quarterly_periods(cutoff))  
  costData[ , ':='(dummy = PostingDate)]
  data.table::setkey(costData, PostingDate, dummy)
  data.table::setkey(timePeriods, origin_period, cutoff_period)

  costDataPeriods <- data.table::foverlaps(x = costData, 
                                           y = timePeriods, 
                                           nomatch = 0L)
  costData[ , ':='(dummy = NULL)]
```

```{r computeSummaries, include=FALSE}
  # Summation per quarter
  costDataSum <- data.table::rbindlist(l = list(costDataPeriods[ , 
                                                          .(Amount = sum(Amount)), 
                                              by = c("category",
                                                     "origin_period",
                                                     "cutoff_period", 
                                                     "period")],
                                    costDataPeriods[ , .(category = "Subtotal",
                                                  Amount = sum(Amount)),
                                              by = c("origin_period",
                                                     "cutoff_period", 
                                                     "period")]),
                                    use.names = TRUE, fill = TRUE)

  incomeDataSum <- data.table::rbindlist(l = list(
    incomeDataPeriods[ , .(Amount = sum(Amount)), 
                by = c("category",
                       "origin_period",
                       "cutoff_period", 
                       "period")],
    incomeDataPeriods[ , .(category = "Subtotal", Amount = sum(Amount)), 
                by = c("origin_period",
                       "cutoff_period", 
                       "period")]),
                                    use.names = TRUE, fill = TRUE)
  # stack income and cost to calculate net profit
  # -- set the sign of the Amount in cost negative to reflect losses.
  stackSubtotals <- 
    data.table::rbindlist(l = list(incomeDataSum[category == "Subtotal"],
                                      costDataSum[category == "Subtotal"
                                                  ][ , c(list(Amount = -1*Amount),
                                                           .SD),
                               .SDcols = setdiff(names(costDataSum),                                                                          "Amount")]),
                               use.names = TRUE, fill = TRUE)
  
  netProfit <- stackSubtotals[, .(category = "Net Profit", 
                                  Amount = sum(Amount)),
                              by = c("origin_period",
                                     "cutoff_period", 
                                     "period")]
  
  # order from smallest to largest so that "Subtotal" is the final entry
  # per quarter
  data.table::setorder(x = costDataSum, 
                       origin_period, cutoff_period, period,
                       Amount
                       )
  data.table::setorder(x = incomeDataSum, 
                       origin_period, cutoff_period, period,
                       Amount
                       )
 mytable <- data.table::rbindlist(l = list(incomeDataSum, 
                                 costDataSum,
                                 netProfit), use.names = TRUE)
  mytable[ , ':='( Amount = scales::dollar_format(negative_parens = TRUE
                                                )(mytable$Amount))]
  
```




# {.tabset}  

## Quarter 1
```{r PnLTable1, include=TRUE, echo=FALSE, class.source="indent", class.output="indent"}

incomeTable(mytable, myperiod = 1)

```


## Quarter 2
```{r PnLTable2, include=TRUE, echo=FALSE, class.source="indent", class.output="indent"}

incomeTable(mytable, myperiod = 2)

```  

## Quarter 3
```{r PnLTable3, include=TRUE, echo=FALSE, class.source="indent", class.output="indent"}

incomeTable(mytable, myperiod = 3)

```


## Quarter 4
```{r PnLTable4, include=TRUE, echo=FALSE, class.source="indent", class.output="indent"}

incomeTable(mytable, myperiod = 4)

```

#

<br>
<br>


